---
title: "Louisiana Mortality Data Description"
author: "Grace Arkfeld"
date: "2024-02-20"
output: pdf_document
---
```{r}
# Set the CRAN mirror
chooseCRANmirror(ind=1)  # Choose the mirror closest to your location

# Install and load necessary packages
install.packages("readr")
library(readr)

# Read the CSV file from Dropbox using the correct link structure
la_mort <- read_csv("https://www.dropbox.com/s/fzsnhfd3lq80v2o3sag6c/la_mort.csv?dl=1")

# Check the first few rows of the dataset to ensure it was loaded correctly
head(la_mort)
```
```{r}
library(dplyr)
```
```{r}
library(readr)
```
```{r}
la_mortality <- read_csv("https://www.dropbox.com/s/fzsnhfd3lq80v2o3sag6c/la_mort.csv?dl=1/LA_Mortality_File.csv")
```
```{r}
# Assuming you've already loaded the dataset into R and named it la_mort

# Get the list of variable names
variable_names <- names(la_mort)
```
## Data Elements Description

1. *restatus*  
   - Description: Resident Status
     - United States Occurrence:
         - State and county of Occurrence and Residence are the same.
         - State of Occurrence and Residence are the same, but County is different.
         - State of Occurrence and Residence are different, but both are in the U.S.
         - State of Occurrence is one of the 50 States or D.C., but Place of Residence is outside the U.S.
     - Puerto Rico Occurrence:
         - Territory and County-equivalent of Occurrence and Residence are the same.
         - Territory of Occurrence and Residence are the same, but County-equivalent is different.
         - Territory of occurrence and residence are different, but both are a Territory.
         - Occurred in Puerto Rico to a resident of any other place.
     - Virgin Islands Occurrence:
         - Territory and County-equivalent of Occurrence and Residence are the same.
         - Territory of Occurrence and Residence are the same, but County-equivalent is different.
         - Territory of occurrence and residence are different, but both are a Territory.
         - Occurred in Virgin Islands to a resident of any other place.

2. *education*  
   - Description: Education Level
     - 2019 Mortality Multiple Cause-of-Death Public Use Record Tape Location Size Field Item and Code Outline:
         - 2 Education (1989 revision):
             - 00: No formal education
             - 01-08: Years of elementary school
             - 09: 1 year of high school
             - 10: 2 years of high school
             - 11: 3 years of high school
             - 12: 4 years of high school
             - 13: 1 year of college
             - 14: 2 years of college
             - 15: 3 years of college
             - 16: 4 years of college
             - 17: 5 or more years of college
             - 99: Not stated
         - 1 Education (2003 revision):
             - Field is blank for registration areas that are using the 1989 revision format of the item.
             - 1: 8th grade or less
             - 2: 9 -12th grade, no diploma
             - 3: high school graduate or GED completed
             - 4: some college credit, but no degree
             - 5: Associate degree
             - 6: Bachelor’s degree
             - 7: Master’s degree
             - 8: Doctorate or professional degree
             - 9: Unknown
     - Education reporting flag:
         - 0: 1989 revision of education item on certificate
         - 1: 2003 revision of education item on certificate
         - 2: No education item on certificate
3. *month_of_death*  
   - Description: Month of Death
     - 2019 Mortality Multiple Cause-of-Death Public Use Record Tape Location Size Field Item and Code Outline:
         - 1: January
         - 2: February
         - 3: March
         - 4: April
         - 5: May
         - 6: June
         - 7: July
         - 8: August
         - 9: September
         - 10: October
         - 11: November
         - 12: December

4. *sex*  
   - Description: Sex
     - 69: M (Male)
     - 70: F (Female)

5. *age*  
   - Description: Age at Death
     - 2019 Mortality Multiple Cause-of-Death Public Use Record Tape Location Size Field Item and Code Outline:
         - 70-73: Detail Age
             - Four positions are used to code detail age.
             - Location 70 identifies age in years, months, days, etc.
             - Locations 71-73 are the number of years, months, days, etc.
         - 74: Age Substitution Flag
             - If reported age is unknown but a valid age is calculated using dates of birth and death, the calculated age is substituted for the unknown reported age.
             - 0: Calculated age is not substituted for reported age
             - 1: Calculated age is substituted for reported age
         - 75-76: Age Recode 52
             - Codes representing different age ranges from under 1 hour to 125 years and over, with categories such as "1 year", "2 years", "3 years", "4 years", "5 - 9 years", "10 - 14 years", etc.
         - 77-78: Age Recode 27
             - Codes representing different age ranges from under 1 month to "Age not stated", with categories such as "Under 1 month", "1 month - 11 months", "1 year", "2 years", "3 years", "4 years", "5 - 9 years", etc.
         - 79-80: Age Recode 12
             - Codes representing different age ranges from under 1 year to "Age not stated", with categories such as "Under 1 year", "1 - 4 years", "5 - 14 years", "15 - 24 years", "25 - 34 years", "35 - 44 years", "45 - 54 years", etc.
         - 81-82: Infant Age Recode
             - Codes representing different age ranges for infants, from under 1 hour to "Age 1 year and over or not stated", with categories such as "Under 1 hour", "1 - 23 hours", "1 day", "2 days", "3 days", "4 days", "5 days", "6 days", "7-13 days", etc.



